---
title: |
  | Accelerometer-based prediction of skeletal mechanical loading during walking in normal weight to severely obese subjects:
  | Data analysis report
author: "Lucas Veras"
date: "2019-12-23"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: 
      collapsed: false
    code_folding: "show"
    highlight: tango
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

Considering the prevalence of obesity [[@Guthold_2018]](#refs), its associated fracture risk [[@Li_2017, @Hind_2018, @Kaze_2018]] and that these patients are especially prone to develop loading associated musculoskeletal injuries [[@Viester_2013]], there is a need to develop mechanical loading prediction equations that are accurate for overweight and obese subjects in order to precisely determine and monitor exercise-associated mechanical loading in these patients. Therefore, the purpose of this study was to develop prediction equations based on accelerometer data able to accurately predict ground reaction forces (GRF) and loading rate (LR) on a broad range of body masses, from normal weight to severely obese subjects, setting thereby, a base for objective prescription and monitoring of exercise mechanical loads.

# Load data

```{r load_packages}
library(knitr)
library(broman)
library(tidyverse)
library(here)
library(nlme)
library(piecewiseSEM)
library(ez)
source(here("R", "cross_validate_mixed_model.R"))
source(here("R", "accuracy_indices.R"))
source(here("R", "get_BA_plot.R"))
```

Initially, 2 files needed to be loaded, each containing data from one of the accelerometer placements utilized in the study: lower back and hip.

```{r load_data}
back <- read_csv(here("data", "back_sec.csv")) %>% 
  select(ID, speed, body_mass, height, BMI, sex, age, pVGRF_N, pVACC_g, pRGRF_N, pRACC_g) %>% 
  filter(speed <= 6)

hip <- read_csv(here("data", "hip_sec.csv")) %>% 
  select(ID, speed, body_mass, height, BMI, sex, age, pVGRF_N, pVACC_g, pRGRF_N, pRACC_g) %>% 
  filter(speed <= 6)
```

A piece of a data frame used in all subsequent analysis can be seen below.

```{r data_example}
back
```

# Mechanical loading X acceleration relationships

We first determined the relationship between peak acceleration (pACC) and peak GRF (pGRF), and between peak ACC acceleration transient rate (pATR) and peak LR (pLR) obtained during the incremental walking speeds used in our experimental protocol. This was performed in order to verify the consistency of these relationships for increasing pACC and pATR values and the ability of the accelerometer to discriminate differences in pGRF and pLR between subjects in different body mass indices (BMI) classes. A scatterplot with these relationships for all two accelerometer placements tested and for both the resultant and its vertical component is depicted in Figure 1. Generally, as expected, pACC recorded by accelerometers in all placements, showed a linear increase as a function of pGRF increases. Also, the recorded ACC were shown to be able to provide a good discrimination between different BMI classes, as for the same registered pACC the pGRF tended to be consistently higher for subjects in higher BMI classes. As was also expected, the pATR recorded by the accelerometers demonstrated a linear increase as the pLR increases. Contrarily to what was observed on the relationship between pACC and pGRF, the pATR x pLR plot does not provide a good discrimination between different BMI classes. Moreover, the values at low magnitudes showed a high degree of clustering, whereas the values at higher magnitudes tended to a greater dispersion. 

```{r GRF_vs_ACC_plot, fig.height=20, fig.width=15, dev='svg'}
source(here("figs", "fig1.R"))

GRF_ACC_plot_grid
```

Code to generate these plots can be seen [here](https://github.com/verasls/walking_GRF_ACC/blob/master/figs/fig1.R).

# Ground reaction force prediction

## Linear mixed models

Linear mixed models (LMM) were developed to predict peak resultant GRF (pRGRF) and peak vertical GRF (pVGRF). Distinct LMMs were developed with data from lowee back and hip accelerometer placement (`back` and `hip` data frames) using the `lme()` function of the **nlme** [package](https://cran.r-project.org/web/packages/nlme/nlme.pdf). Covariance structure used was an autoregressive process of order 1 (`correlation = corAR1`) and maximum likelihood method was used for estimating parameters (`method = "ML"`). Predictors tested on pRGRF models were body mass and peak resultant acceleration (pRACC), while on pVGRF were body mass and vertical acceleration (pVACC). All of them were tested as fixed effects and have shown to be significant predictors (*e.g.:* `fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass`). Both random intercept and slopes were tested, but only the random intercept inclusion has showed models improvement (`random = ~ 1 | ID`). Linear, quadratic and cubic polynomial simulations were also tested, whereas the last one did not contribute significantly to the models. Final models were chosen according to -2 log-likelihood statistics. Traditional coefficient of determination (R^2^) was represented by conditional R^2^ [[@naka13]](#refs) computed with `rsquared()` function of the **piecewiseSEM** [package](http://jslefche.github.io/piecewiseSEM/articles/piecewiseSEM.html), that estimates the variance explained by the whole model [[@r216]](#refs).

Code to build the LMMs, as well as their summary and R^2^, can be found below.

```{r LMM}
# For resultant peak ground reaction force
back_res_LMM <- lme(
  fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass + pRACC_g : body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = back
)
r2_back_res_LMM <- rsquared(back_res_LMM)

summary(back_res_LMM)
r2_back_res_LMM

hip_res_LMM <- lme(
  fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass + pRACC_g : body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = hip
)
r2_hip_res_LMM <- rsquared(hip_res_LMM)


summary(hip_res_LMM)
r2_hip_res_LMM

# For vertical peak ground reaction force
back_vert_LMM <- lme(
  fixed = pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass + pVACC_g : body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = back
)
r2_back_vert_LMM <- rsquared(back_vert_LMM)

summary(back_vert_LMM)
r2_back_vert_LMM

hip_vert_LMM <- lme(
  fixed = pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass + pVACC_g : body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = hip
)
r2_hip_vert_LMM <- rsquared(hip_vert_LMM)

summary(hip_vert_LMM)
r2_hip_vert_LMM
```

```{r R2_range, echo=FALSE}
minR2 <- myround(min(c(r2_back_res_LMM[[6]], r2_hip_res_LMM[[6]], r2_back_vert_LMM[[6]], r2_hip_vert_LMM[[6]])), digits = 2)
maxR2 <- myround(max(c(r2_back_res_LMM[[6]], r2_hip_res_LMM[[6]], r2_back_vert_LMM[[6]], r2_hip_vert_LMM[[6]])), digits = 2)
```

In all models, conditional R^2^ values ranged from `r minR2` to `r maxR2`.

## Validation analysis

Model validation was assessed by the leave-one-out cross-validation (LOOCV) method. This approach is recommended when a different sample is not available for cross-validation and it provides an insight on the model potential to predict outcomes in a new independent sample [[@staud12]](#refs). For LOOCV each participant’s data was separated into a testing dataset (one at a time) with the remaining data being in the training dataset. New LMM, with the same outcomes and predictors as determined for the entire sample, were developed using the training dataset and then used to predict pGRF for the only participant in the testing dataset. This process was repeated for all participants (64 times). Data from testing dataset were used in the remaining statistical analysis.

LOOCV of the LMM was done with the `cross_validate_mixed_model()` [function](https://github.com/verasls/walking_GRF_ACC/blob/master/R/cross_validate_mixed_model.R) and is shown below.

```{r LOOCV}
# For resultant peak ground reaction force
fix_eff    <- pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass
rand_eff   <- ~ 1 | ID
# Back
LOOCV_back_res_LMM <- do.call(rbind, (lapply(unique(back$ID), cross_validate_mixed_model, df = back)))
# Hip
LOOCV_hip_res_LMM <- do.call(rbind, (lapply(unique(hip$ID), cross_validate_mixed_model, df = hip)))

# For vertical peak ground reaction force
fix_eff    <- pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass
rand_eff   <- ~ 1 | ID
# Back
LOOCV_back_vert_LMM <- do.call(rbind, (lapply(unique(back$ID), cross_validate_mixed_model, df = back)))
# Hip
LOOCV_hip_vert_LMM <- do.call(rbind, (lapply(unique(hip$ID), cross_validate_mixed_model, df = hip)))
```

### Bland-Altman plots

Bland-Altman plots were used to examine the agreement between pGRF measured with force plates and those predicted through the regression equations. The difference of the actual and predicted pGRF was plotted against their mean. Bias was expressed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted pGRF [[@ba86]](#refs).

Bland-Altman plots can be seen below for all accelerometer placements. Panel **A** shows plots of pRGRF and panel **B** shows plots of pVGRF.

```{r BA_plots, fig.height=20, fig.width=15, dev='svg'}
source(here("figs", "fig2.R"))

BA_plot_grid
```

Code to generate these plots can be seen [here](https://github.com/verasls/walking_GRF_ACC/blob/master/figs/fig2.R).

<br>

One-sample t-tests were run to check whether bias from each accelerometer placement, in both pRGRF and pVGRF, was significantly different than 0. These tests were run using the `t.test()` function of the base R with the argument `mu = 0`. No significant differences were found.

```{r t_test}
# For resultant peak ground reaction force
# Back
LOOCV_back_res_LMM$diff <- LOOCV_back_res_LMM$pRGRF_N - LOOCV_back_res_LMM$pRGRF_N_predicted
LOOCV_back_res_LMM$mean <- (LOOCV_back_res_LMM$pRGRF_N + LOOCV_back_res_LMM$pRGRF_N_predicted) / 2
t.test(LOOCV_back_res_LMM$diff, mu = 0)
# Hip
LOOCV_hip_res_LMM$diff <- LOOCV_hip_res_LMM$pRGRF_N - LOOCV_hip_res_LMM$pRGRF_N_predicted
LOOCV_hip_res_LMM$mean <- (LOOCV_hip_res_LMM$pRGRF_N + LOOCV_hip_res_LMM$pRGRF_N_predicted) / 2
t.test(LOOCV_hip_res_LMM$diff, mu = 0)

# For vertical peak ground reaction force
# Back
LOOCV_back_vert_LMM$diff <- LOOCV_back_vert_LMM$pVGRF_N - LOOCV_back_vert_LMM$pVGRF_N_predicted
LOOCV_back_vert_LMM$mean <- (LOOCV_back_vert_LMM$pVGRF_N + LOOCV_back_vert_LMM$pVGRF_N_predicted) / 2
t.test(LOOCV_back_vert_LMM$diff, mu = 0)
# Hip
LOOCV_hip_vert_LMM$diff <- LOOCV_hip_vert_LMM$pVGRF_N - LOOCV_hip_vert_LMM$pVGRF_N_predicted
LOOCV_hip_vert_LMM$mean <- (LOOCV_hip_vert_LMM$pVGRF_N + LOOCV_hip_vert_LMM$pVGRF_N_predicted) / 2
t.test(LOOCV_hip_vert_LMM$diff, mu = 0)
```

Also, linear regressions were applied to identify if there was any proportional bias, that is, if bias was related with the magnitude of the mean between measured and predicted pGRF [[@giav15]](#refs). Linear regressions were run using the `lm()` function of the base R.

Results showed a significant proportional bias (*p* < 0.05), however, with a low magnitude (highest R^2^ = 0.032). These results suggest that despite there is a trend for underestimation at increasingly higher pGRF values, the magnitude of this effect is neglectable.

```{r BA_LR, fig.height=15, fig.width=13, dev='svg'}
# For resultant peak ground reaction force
# Back
back_res_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_back_res_LMM)
summary(back_res_BA_plot_LR)
# Hip
hip_res_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_hip_res_LMM)
summary(hip_res_BA_plot_LR)

# For vertical peak ground reaction force
# Back
back_vert_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_back_vert_LMM)
summary(back_vert_BA_plot_LR)
# Hip
hip_vert_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_hip_vert_LMM)
summary(hip_vert_BA_plot_LR)
```

### Indices of accuracy

To evaluate models prediction accuracy, mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE) were computed These indices were computed with the `accuracy_indices()` [function](https://github.com/verasls/walking_GRF_ACC/blob/master/R/accuracy_indices.R).

```{r accuracy_indices}
# For resultant peak ground reaction force
# Back
back_res_accuracy <- accuracy_indices(LOOCV_back_res_LMM, "pRGRF_N", "pRGRF_N_predicted")
# Hip
hip_res_accuracy <- accuracy_indices(LOOCV_hip_res_LMM, "pRGRF_N", "pRGRF_N_predicted")

# For vertical peak ground reaction force
# Back
back_vert_accuracy <- accuracy_indices(LOOCV_back_vert_LMM, "pVGRF_N", "pVGRF_N_predicted")
# Hip
hip_vert_accuracy <- accuracy_indices(LOOCV_hip_vert_LMM, "pVGRF_N", "pVGRF_N_predicted")
```

Table below shows these indices for each accelerometer placement in both pRGRF and pVGRF.

```{r accuracy_table}
accuracy <- data.frame(
  acc_placement = c(rep("Back", 2), rep("Hip", 2)),
  condition     = rep(c("pRGRF", "pVGRF"), 2),
  MAE           = c(
    round(back_res_accuracy[[5]], 1), round(back_vert_accuracy[[5]], 1), 
    round(hip_res_accuracy[[5]], 1), round(hip_vert_accuracy[[5]], 1)
    ),
  MAPE          = c(
    paste(round(back_res_accuracy[[12]], 1), "%", sep = ""), paste(round(back_vert_accuracy[[12]], 1), "%", sep = ""), 
    paste(round(hip_res_accuracy[[12]], 1), "%", sep = ""), paste(round(hip_vert_accuracy[[12]], 1), "%", sep = "")
    ),
  RMSE          = c(
    round(back_res_accuracy[[7]], 1), round(back_vert_accuracy[[7]], 1), 
    round(hip_res_accuracy[[7]], 1), round(hip_vert_accuracy[[7]], 1)
    )
)
kable(accuracy, col.names = c("Accelerometer Placement", "GRF", "MAE", "MAPE", "RMSE"), align = c("l", "l", "r", "r", "r"))
```

All conditions showed a MAE from near 65 N to near 81 N, MAPE from near 6% to near 7% and RMSE from near 84 N to 84 N.

### ANOVA

A series of repeated measures analysis of variance (ANOVA) were run to assess whether pGRF predicted from the regression equations were significantly different from those measured with FP. Walking speeds, accelerometer placements (lower back and hip), and the interaction effect (speed X accelerometer placements) were considered for analysis. These procedures were taken separately for resultant and its vertical component. If assumptions of sphericity were violated (*p* < 0.05), the conservative Greenhouse–Geisser correction factor would be applied to adjust the degrees of freedom. Post-hoc analyses would be conducted using pairwise comparisons with Holm's correction if a significant difference were observed among actual and predicted pGRF.

In order to run the repeated measures ANOVA, new data frames needed to be built, for both pRGRF and pVGRF, putting all the pGRF values in a single variable and grouping then by accelerometer placement or actual value in another variable. An example of code to tidy the data is shown below.

```{r build_pRGRF_df}
## Predicted pRGRF
back_res_pred <- LOOCV_back_res_LMM %>% 
  select(ID, speed, pRGRF_N_predicted) %>% 
  spread(key = speed, value = pRGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = back
  )

hip_res_pred <- LOOCV_hip_res_LMM %>% 
  select(ID, speed, pRGRF_N_predicted) %>% 
  spread(key = speed, value = pRGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = hip
  )

res_pred_df <- back_res_pred %>% 
  full_join(hip_res_pred, by = c("ID", "speed")) %>% 
  na.omit()

## Actual pRGRF
back_res_actual <- LOOCV_back_res_LMM %>% 
  select(ID, speed, pRGRF_N) %>% 
  spread(key = speed, value = pRGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_back
  )

hip_res_actual <- LOOCV_hip_res_LMM %>% 
  select(ID, speed, pRGRF_N) %>% 
  spread(key = speed, value = pRGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_hip
  )

res_actual_df <- back_res_actual %>% 
  full_join(hip_res_actual, by = c("ID", "speed")) %>% 
  na.omit()

res_actual_df <- res_actual_df %>% 
  mutate(actual = (actual_back + actual_hip) / 2) %>% 
  select(ID, speed, actual) 

## Merge predicted and actual data frames
res_ANOVA_df <- res_pred_df %>% 
  full_join(res_actual_df, by = c("ID", "speed")) %>% 
  gather(
    back, hip, actual,
    key = "group",
    value = "pRGRF"
  )
res_ANOVA_df$ID    <- as.factor(res_ANOVA_df$ID)
res_ANOVA_df$speed <- as.factor(res_ANOVA_df$speed)
res_ANOVA_df$group <- as.factor(res_ANOVA_df$group)
```

```{r build_pVGRF_df, echo=FALSE}
## Predicted pVGRF
back_vert_pred <- LOOCV_back_vert_LMM %>% 
  select(ID, speed, pVGRF_N_predicted) %>% 
  spread(key = speed, value = pVGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = back
  )

hip_vert_pred <- LOOCV_hip_vert_LMM %>% 
  select(ID, speed, pVGRF_N_predicted) %>% 
  spread(key = speed, value = pVGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = hip
  )

vert_pred_df <- back_vert_pred %>% 
  full_join(hip_vert_pred, by = c("ID", "speed")) %>% 
  na.omit()

## Actual pVGRF
back_vert_actual <- LOOCV_back_vert_LMM %>% 
  select(ID, speed, pVGRF_N) %>% 
  spread(key = speed, value = pVGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_back
  )

hip_vert_actual <- LOOCV_hip_vert_LMM %>% 
  select(ID, speed, pVGRF_N) %>% 
  spread(key = speed, value = pVGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_hip
  )

vert_actual_df <- back_vert_actual %>% 
  full_join(hip_vert_actual, by = c("ID", "speed")) %>% 
  na.omit()

vert_actual_df <- vert_actual_df %>% 
  mutate(actual = (actual_back + actual_hip) / 2) %>% 
  select(ID, speed, actual) 

## Merge predicted and actual data frames
vert_ANOVA_df <- vert_pred_df %>% 
  full_join(vert_actual_df, by = c("ID", "speed")) %>% 
  gather(
    back, hip, actual,
    key = "group",
    value = "pVGRF"
  )
vert_ANOVA_df$ID    <- as.factor(vert_ANOVA_df$ID)
vert_ANOVA_df$speed <- as.factor(vert_ANOVA_df$speed)
vert_ANOVA_df$group <- as.factor(vert_ANOVA_df$group)
```

Repeated measures ANOVAs were performed using the `ezANOVA()` function of the **ez** [package](https://cran.r-project.org/web/packages/ez/ez.pdf). Summary statistics for pRGRF and pVGRF ANOVA can be seen below, respectively.

```{r ANOVA}
# For resultant peak ground reaction force
res_ANOVA <- ezANOVA(
  data     = res_ANOVA_df,
  dv       = pRGRF,
  wid      = ID,
  within   = .(speed, group),
  detailed = TRUE,
  type     = 3
)
res_ANOVA

# For vertical peak ground reaction force
vert_ANOVA <- ezANOVA(
  data     = vert_ANOVA_df,
  dv       = pVGRF,
  wid      = ID,
  within   = .(speed, group),
  detailed = TRUE,
  type     = 3
)
vert_ANOVA
```

For either resultant and its vertical component, repeated measures ANOVA demonstrated that actual and predicted pGRF increased significantly (*p* < 0.001) along with increments in walking speed. 

<br>

To assess for any significant differences between actual and predicted pGRF in each speed, five separate repeated measures ANOVA were done, one for each walking speed.

```{r}
# For resultant peak ground reaction force
# 2 km/h
res_ANOVA_s2 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 2),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 3 km/h
res_ANOVA_s3 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 3),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 4 km/h
res_ANOVA_s4 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 4),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 5 km/h
res_ANOVA_s5 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 5),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 6 km/h
res_ANOVA_s6 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 6),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# For vertical peak ground reaction force
# 2 km/h
vert_ANOVA_s2 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 2),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 3 km/h
vert_ANOVA_s3 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 3),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 4 km/h
vert_ANOVA_s4 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 4),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 5 km/h
vert_ANOVA_s5 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 5),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 6 km/h
vert_ANOVA_s6 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 6),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)
```

No significant differences (*p* > 0.05) between actual and predicted pGRF in each speed were found.

These results can be seen in figure below.

```{r ANOVA_GRF_plot, fig.height=5, fig.width=10, dev='svg'}
source(here("figs", "fig3.R"))

GRF_plot_grid
```

## Regressions comparison

Our equation was compared with a previously published reference equation [[@Neugebauer_2014]](#refs), in which Neugebauer et al. used a similar approach for the prediction of pGRF. Comparison was performed using the equation to predict pVGRF from hip-worn accelerometers, as it was the only suitable with our data. The analysis was performed in three ways using: i) a subsample of participants with normal weight or overweight (BMI ≥ 18.5 and < 30 kg.m-2); ii) a subsample of obese participants (BMI ≥ 30 kg.m-2); iii) the whole sample. 

New data frames needed to be built for each of the subsamples and the whole sample. Then, pVGRF was predicted using the reference equation. Code to build the data frames is shown below.

```{r build_comparison_df}
# Prepare  data frames
non_obese_df <- read_csv("~/Dropbox/Projects/walking_GRF_ACC/LOOCV_hip_vert.csv") %>% 
  select(ID, speed, body_mass, BMI, BMI_cat, pVACC_g, pVGRF_N, pVGRF_N_predicted) %>% 
  filter(BMI < 30)

obese_df <- read_csv("~/Dropbox/Projects/walking_GRF_ACC/LOOCV_hip_vert.csv") %>% 
  select(ID, speed, body_mass, BMI, BMI_cat, pVACC_g, pVGRF_N, pVGRF_N_predicted) %>% 
  filter(BMI >= 30)

whole_sample_df <- read_csv("~/Dropbox/Projects/walking_GRF_ACC/LOOCV_hip_vert.csv") %>% 
  select(ID, speed, body_mass, BMI, BMI_cat, pVACC_g, pVGRF_N, pVGRF_N_predicted)

# Apply Neugebauer 2014 equation
non_obese_df$pVGRF_N_Neugebauer <- NA
for (i in 1:nrow(non_obese_df)) {
  non_obese_df$pVGRF_N_Neugebauer[i] <-  
    exp(5.247 + (0.271 * non_obese_df$pVACC_g[i]) + (0.014 * non_obese_df$body_mass[i]))
}

obese_df$pVGRF_N_Neugebauer <- NA
for (i in 1:nrow(obese_df)) {
  obese_df$pVGRF_N_Neugebauer[i] <-  
    exp(5.247 + (0.271 * obese_df$pVACC_g[i]) + (0.014 * obese_df$body_mass[i]))
}

whole_sample_df$pVGRF_N_Neugebauer <- NA
for (i in 1:nrow(whole_sample_df)) {
  whole_sample_df$pVGRF_N_Neugebauer[i] <-  
    exp(5.247 + (0.271 * whole_sample_df$pVACC_g[i]) + (0.014 * whole_sample_df$body_mass[i]))
}
```

Bland-Altman plots were used to confront the agreement between pVGRF measured with FP and those predicted using both regression equations. 

```{r comparison_BA_plots, fig.height=15, fig.width=13, dev='svg'}
source(here("figs", "fig4.R"))

BA_plot_grid
```

Code to generate these plots can be seen [here](https://github.com/verasls/walking_GRF_ACC/blob/master/figs/fig4.R).

A lower dispersion around the bias value, as well as a lower percentage of values falling off the LoA can be appreciated in our equation. Additionally, while the bias in our equation tended to be zero, in the reference equation bias was always > 0 (*p* < 0.05), showing a consistent underestimation of pVGRF.

<br>

Also, to assess the prediction accuracy, MAE, MAPE and RMSE were calculated using pVGRF values predicted from both equations.

```{r comparison_accuracy_indices}
non_obese_our_accuracy     <- accuracy_indices(non_obese_df, "pVGRF_N", "pVGRF_N_predicted")
non_obese_Neug_accuracy    <- accuracy_indices(non_obese_df, "pVGRF_N", "pVGRF_N_Neugebauer")
obese_our_accuracy         <- accuracy_indices(obese_df, "pVGRF_N", "pVGRF_N_predicted")
obese_Neug_accuracy        <- accuracy_indices(obese_df, "pVGRF_N", "pVGRF_N_Neugebauer")
whole_sample_our_accuracy  <- accuracy_indices(whole_sample_df, "pVGRF_N", "pVGRF_N_predicted")
whole_sample_Neug_accuracy <- accuracy_indices(whole_sample_df, "pVGRF_N", "pVGRF_N_Neugebauer")
```


The accuracy indices from our equation were all substantially lower compared to the reference equation indices, with an overall MAPE approximately 3 times smaller for our equation. For both equations, the MAPE was lower in the obese subsample than in the non-obese subsample

# R session info

```{r session_info}
devtools::session_info()
```

# References {#refs}