---
title: |
  | Prediction of walking ground reaction forces based on accelerometry: 
  | Data analysis report
author: "Lucas Veras"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: 
      collapsed: false
    code_folding: "show"
    highlight: tango
editor_options: 
  chunk_output_type: console
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Background

Ground reaction force (GRF) is typically measured by force plates (FP), being these instruments usually limited to laboratory settings. Accelerometers may be a viable device to predict peak GRF (pGRF) in field conditions. Nevertheless, few studies have tested its validity. Therefore, the purpose of this study was to develop regression equations to predict peak resultant GRF and its peak vertical component based on ankle, back and hip accelerometry data during walking.

# Load data

```{r load_packages}
library(knitr)
library(broman)
library(tidyverse)
library(here)
library(nlme)
library(piecewiseSEM)
library(ez)
source(here("R", "cross_validate_mixed_model.R"))
source(here("R", "accuracy_indices.R"))
source(here("R", "get_BA_plot.R"))
```

Initially, 3 files needed to be loaded, each containing data from one of the accelerometer placements utilized in the study: ankle, back and hip.

```{r load_data}
ankle <- read_csv(here("data", "ankle_sec.csv")) %>% 
  select(ID, speed, body_mass, height, BMI, sex, age, pVGRF_N, pVACC_g, pRGRF_N, pRACC_g) %>% 
  filter(speed <= 6)

back <- read_csv(here("data", "back_sec.csv")) %>% 
  select(ID, speed, body_mass, height, BMI, sex, age, pVGRF_N, pVACC_g, pRGRF_N, pRACC_g) %>% 
  filter(speed <= 6)

hip <- read_csv(here("data", "hip_sec.csv")) %>% 
  select(ID, speed, body_mass, height, BMI, sex, age, pVGRF_N, pVACC_g, pRGRF_N, pRACC_g) %>% 
  filter(speed <= 6)
```

A piece of a data frame used in all subsequent analysis can be seen below.

```{r data_example}
ankle
```

# Linear mixed models

Linear mixed models (LMM) were used to estimate peak resultant ground reaction force (pRGRF) and peak vertical ground reaction force (pVGRF). Distinct LMMs were developed with data from ankle, back and hip accelerometer placement (`ankle`, `back` and `hip` data frames) using the `lme()` function of the **nlme** [package](https://cran.r-project.org/web/packages/nlme/nlme.pdf). Covariance structure used was an autoregressive process of order 1 (`correlation = corAR1`) and maximum likelihood method was used for estimating parameters (`method = "ML"`). Predictors tested on pRGRF models were body mass and peak resultant acceleration (pRACC), while on pVGRF were body mass and vertical acceleration (pVACC). All of them were tested as fixed effects and have shown to be significant predictors (*e.g.:* `fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass`). Both random intercept and slopes were tested, but only the random intercept inclusion has showed models improvement (`random = ~ 1 | ID`). Linear, quadratic and cubic polynomial simulations were also tested, whereas the last one did not contribute significantly to the models. Final models were chosen according to -2 log-likelihood statistics. Traditional coefficient of determination (R^2^) was represented by conditional R^2^ [[@naka13]](#refs) computed with `rsquared()` function of the **piecewiseSEM** [package](http://jslefche.github.io/piecewiseSEM/articles/piecewiseSEM.html), that estimates the variance explained by the whole model [[@r216]](#refs).

Code to build the LMMs, as well as their summary and R^2^, can be found below.

```{r LMM}
# For resultant peak ground reaction force
ankle_res_LMM <- lme(
  fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = ankle
)
r2_ankle_res_LMM <- rsquared(ankle_res_LMM)

summary(ankle_res_LMM)
r2_ankle_res_LMM

back_res_LMM <- lme(
  fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = back
)
r2_back_res_LMM <- rsquared(back_res_LMM)

summary(back_res_LMM)
r2_back_res_LMM

hip_res_LMM <- lme(
  fixed = pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = hip
)
r2_hip_res_LMM <- rsquared(hip_res_LMM)


summary(hip_res_LMM)
r2_hip_res_LMM

# For vertical peak ground reaction force
ankle_vert_LMM <- lme(
  fixed = pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = ankle
)
r2_ankle_vert_LMM <- rsquared(ankle_vert_LMM)

summary(ankle_vert_LMM)
r2_ankle_vert_LMM

back_vert_LMM <- lme(
  fixed = pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = back
)
r2_back_vert_LMM <- rsquared(back_vert_LMM)

summary(back_vert_LMM)
r2_back_vert_LMM

hip_vert_LMM <- lme(
  fixed = pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass,
  random = ~ 1 | ID,
  method = "ML",
  correlation = corAR1(),
  data = hip
)
r2_hip_vert_LMM <- rsquared(hip_vert_LMM)

summary(hip_vert_LMM)
r2_hip_vert_LMM
```

```{r R2_range, echo=FALSE}
minR2 <- myround(min(c(r2_ankle_res_LMM[[6]], r2_back_res_LMM[[6]], r2_hip_res_LMM[[6]], r2_ankle_vert_LMM[[6]], r2_back_vert_LMM[[6]], r2_hip_vert_LMM[[6]])), digits = 2)
maxR2 <- myround(max(c(r2_ankle_res_LMM[[6]], r2_back_res_LMM[[6]], r2_hip_res_LMM[[6]], r2_ankle_vert_LMM[[6]], r2_back_vert_LMM[[6]], r2_hip_vert_LMM[[6]])), digits = 2)
```

In all models, conditional R^2^ values ranged from `r minR2` to `r maxR2`.

# Validation analysis

Model validation was assessed by the leave-one-out cross-validation (LOOCV) method. This approach is recommended when a different sample is not available for cross-validation and it provides an insight on the model potential to predict outcomes in a new independent sample [[@staud12]](#refs). For LOOCV each participant’s data was separated into a testing dataset (one at a time) with the remaining data being in the training dataset. New LMM, with the same outcomes and predictors as determined for the entire sample, were developed using the training dataset and then used to predict pGRF for the only participant in the testing dataset. This process was repeated for all participants (64 times). Data from testing dataset were used in the remaining statistical analysis.

LOOCV of the LMM was done with the `cross_validate_mixed_model()` [function](https://github.com/verasls/walking_GRF_ACC/blob/master/R/cross_validate_mixed_model.R) and is shown below.

```{r LOOCV}
# For resultant peak ground reaction force
fix_eff    <- pRGRF_N ~ pRACC_g + I(pRACC_g^2) + body_mass
rand_eff   <- ~ 1 | ID
# Ankle
LOOCV_ankle_res_LMM <- do.call(rbind, (lapply(unique(ankle$ID), cross_validate_mixed_model, df = ankle)))
# Back
LOOCV_back_res_LMM <- do.call(rbind, (lapply(unique(back$ID), cross_validate_mixed_model, df = back)))
# Hip
LOOCV_hip_res_LMM <- do.call(rbind, (lapply(unique(hip$ID), cross_validate_mixed_model, df = hip)))

# For vertical peak ground reaction force
fix_eff    <- pVGRF_N ~ pVACC_g + I(pVACC_g^2) + body_mass
rand_eff   <- ~ 1 | ID
# Ankle
LOOCV_ankle_vert_LMM <- do.call(rbind, (lapply(unique(ankle$ID), cross_validate_mixed_model, df = ankle)))
# Back
LOOCV_back_vert_LMM <- do.call(rbind, (lapply(unique(back$ID), cross_validate_mixed_model, df = back)))
# Hip
LOOCV_hip_vert_LMM <- do.call(rbind, (lapply(unique(hip$ID), cross_validate_mixed_model, df = hip)))
```

## Bland-Altman plots

Bland-Altman plots were used to examine the agreement between pGRF measured with force plates and those predicted through the regression equations. The difference of the actual and predicted pGRF was plotted against their mean. Bias was expressed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted pGRF [[@ba86]](#refs).

Bland-Altman plots can be seen below for all accelerometer placements. Panel **A** shows plots of pRGRF and panel **B** shows plots of pVGRF.

```{r BA_plots, fig.height=15, fig.width=13, dev='svg'}
source(here("figs", "fig1.R"))

BA_plot_grid
```

Code to generate these plots can be seen [here](https://github.com/verasls/walking_GRF_ACC/blob/master/figs/fig1.R).

<br>

One-sample t-tests were run to check whether bias from each accelerometer placement, in both pRGRF and pVGRF, was significantly different than 0. These tests were run using the `t.test()` function of the base R with the argument `mu = 0`. No significant differences were found.

```{r t_test}
# For resultant peak ground reaction force
# Ankle
LOOCV_ankle_res_LMM$diff <- LOOCV_ankle_res_LMM$pRGRF_N - LOOCV_ankle_res_LMM$pRGRF_N_predicted
LOOCV_ankle_res_LMM$mean <- (LOOCV_ankle_res_LMM$pRGRF_N + LOOCV_ankle_res_LMM$pRGRF_N_predicted) / 2
t.test(LOOCV_ankle_res_LMM$diff, mu = 0)
# Back
LOOCV_back_res_LMM$diff <- LOOCV_back_res_LMM$pRGRF_N - LOOCV_back_res_LMM$pRGRF_N_predicted
LOOCV_back_res_LMM$mean <- (LOOCV_back_res_LMM$pRGRF_N + LOOCV_back_res_LMM$pRGRF_N_predicted) / 2
t.test(LOOCV_back_res_LMM$diff, mu = 0)
# Hip
LOOCV_hip_res_LMM$diff <- LOOCV_hip_res_LMM$pRGRF_N - LOOCV_hip_res_LMM$pRGRF_N_predicted
LOOCV_hip_res_LMM$mean <- (LOOCV_hip_res_LMM$pRGRF_N + LOOCV_hip_res_LMM$pRGRF_N_predicted) / 2
t.test(LOOCV_hip_res_LMM$diff, mu = 0)

# For vertical peak ground reaction force
# Ankle
LOOCV_ankle_vert_LMM$diff <- LOOCV_ankle_vert_LMM$pVGRF_N - LOOCV_ankle_vert_LMM$pVGRF_N_predicted
LOOCV_ankle_vert_LMM$mean <- (LOOCV_ankle_vert_LMM$pVGRF_N + LOOCV_ankle_vert_LMM$pVGRF_N_predicted) / 2
t.test(LOOCV_ankle_vert_LMM$diff, mu = 0)
# Back
LOOCV_back_vert_LMM$diff <- LOOCV_back_vert_LMM$pVGRF_N - LOOCV_back_vert_LMM$pVGRF_N_predicted
LOOCV_back_vert_LMM$mean <- (LOOCV_back_vert_LMM$pVGRF_N + LOOCV_back_vert_LMM$pVGRF_N_predicted) / 2
t.test(LOOCV_back_vert_LMM$diff, mu = 0)
# Hip
LOOCV_hip_vert_LMM$diff <- LOOCV_hip_vert_LMM$pVGRF_N - LOOCV_hip_vert_LMM$pVGRF_N_predicted
LOOCV_hip_vert_LMM$mean <- (LOOCV_hip_vert_LMM$pVGRF_N + LOOCV_hip_vert_LMM$pVGRF_N_predicted) / 2
t.test(LOOCV_hip_vert_LMM$diff, mu = 0)
```

Also, linear regressions were applied to identify if there was any proportional bias, that is, if bias was related with the magnitude of the mean between measured and predicted pGRF [[@giav15]](#refs). Linear regressions were run using the `lm()` function of the base R.

In back and hip placements, linear regression showed a proportional bias (*p* < 0.05), however, though significant (*p* < 0.05), associations magnitude was low (R^2^ < 0.025). The same did not occur to ankle placement (*p* > 0.05).

```{r BA_LR, fig.height=15, fig.width=13, dev='svg'}
# For resultant peak ground reaction force
# Ankle
ankle_res_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_ankle_res_LMM)
summary(ankle_res_BA_plot_LR)
# Back
back_res_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_back_res_LMM)
summary(back_res_BA_plot_LR)
# Hip
hip_res_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_hip_res_LMM)
summary(hip_res_BA_plot_LR)

# For vertical peak ground reaction force
# Ankle
ankle_vert_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_ankle_vert_LMM)
summary(ankle_vert_BA_plot_LR)
# Back
back_vert_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_back_vert_LMM)
summary(back_vert_BA_plot_LR)
# Hip
hip_vert_BA_plot_LR <- lm(diff ~ mean, data = LOOCV_hip_vert_LMM)
summary(hip_vert_BA_plot_LR)
```

## Indices of accuracy

To evaluate models prediction accuracy, mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE) were computed These indices were computed with the `accuracy_indices()` [function](https://github.com/verasls/walking_GRF_ACC/blob/master/R/accuracy_indices.R).

```{r accuracy_indices}
# For resultant peak ground reaction force
# Ankle
ankle_res_accuracy <- accuracy_indices(LOOCV_ankle_res_LMM, "pRGRF_N", "pRGRF_N_predicted")
# Back
back_res_accuracy <- accuracy_indices(LOOCV_back_res_LMM, "pRGRF_N", "pRGRF_N_predicted")
# Hip
hip_res_accuracy <- accuracy_indices(LOOCV_hip_res_LMM, "pRGRF_N", "pRGRF_N_predicted")

# For vertical peak ground reaction force
# Ankle
ankle_vert_accuracy <- accuracy_indices(LOOCV_ankle_vert_LMM, "pVGRF_N", "pVGRF_N_predicted")
# Back
back_vert_accuracy <- accuracy_indices(LOOCV_back_vert_LMM, "pVGRF_N", "pVGRF_N_predicted")
# Hip
hip_vert_accuracy <- accuracy_indices(LOOCV_hip_vert_LMM, "pVGRF_N", "pVGRF_N_predicted")
```

Table below shows these indices for each accelerometer placement in both pRGRF and pVGRF.

```{r accuracy_table}
accuracy <- data.frame(
  acc_placement = c(rep("Ankle", 2), rep("Back", 2), rep("Hip", 2)),
  condition     = rep(c("pRGRF", "pVGRF"), 3),
  MAE           = c(
    round(ankle_res_accuracy[[5]], 1), round(ankle_vert_accuracy[[5]], 1), 
    round(back_res_accuracy[[5]], 1), round(back_vert_accuracy[[5]], 1), 
    round(hip_res_accuracy[[5]], 1), round(hip_vert_accuracy[[5]], 1)
    ),
  MAPE          = c(
    paste(round(ankle_res_accuracy[[12]], 1), ".0%", sep = ""), paste(round(ankle_vert_accuracy[[12]], 1), "%", sep = ""), 
    paste(round(back_res_accuracy[[12]], 1), "%", sep = ""), paste(round(back_vert_accuracy[[12]], 1), "%", sep = ""), 
    paste(round(hip_res_accuracy[[12]], 1), "%", sep = ""), paste(round(hip_vert_accuracy[[12]], 1), "%", sep = "")
    ),
  RMSE          = c(
    round(ankle_res_accuracy[[7]], 1), round(ankle_vert_accuracy[[7]], 1), 
    round(back_res_accuracy[[7]], 1), round(back_vert_accuracy[[7]], 1), 
    round(hip_res_accuracy[[7]], 1), round(hip_vert_accuracy[[7]], 1)
    )
)
kable(accuracy, col.names = c("Accelerometer Placement", "GRF", "MAE", "MAPE", "RMSE"), align = c("l", "l", "r", "r", "r"))
```

All conditions showed a MAE from near 65 N to near 81 N, MAPE from near 6% to near 7% and RMSE from near 84 N to 84 N.

## ANOVA

A series of repeated measures analysis of variance (ANOVA) were run to assess whether pGRF predicted from the regression equations were significantly different from those measured with FP. Walking speeds, accelerometer placements (ankle, back and hip), and the interaction effect (speed X accelerometer placements) were considered for analysis. These procedures were taken separately for resultant and its vertical component. If assumptions of sphericity were violated (*p* < 0.05), the conservative Greenhouse–Geisser correction factor would be applied to adjust the degrees of freedom. Post-hoc analyses would be conducted using pairwise comparisons with Holm's test if a significant difference were observed among actual and predicted pGRF.

In order to run the repeated measures ANOVA, new data frames needed to be built, for both pRGRF and pVGRF, putting all the pGRF values in a single variable and grouping then by accelerometer placement or actual value in another variable. An example of code to tidy the data is shown below.

```{r build_pRGRF_df}
## Predicted pRGRF
ankle_res_pred <- LOOCV_ankle_res_LMM %>% 
  select(ID, speed, pRGRF_N_predicted) %>% 
  spread(key = speed, value = pRGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = ankle
  )

back_res_pred <- LOOCV_back_res_LMM %>% 
  select(ID, speed, pRGRF_N_predicted) %>% 
  spread(key = speed, value = pRGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = back
  )

hip_res_pred <- LOOCV_hip_res_LMM %>% 
  select(ID, speed, pRGRF_N_predicted) %>% 
  spread(key = speed, value = pRGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = hip
  )

res_pred_df <- ankle_res_pred %>% 
  full_join(back_res_pred, by = c("ID", "speed")) %>% 
  full_join(hip_res_pred, by = c("ID", "speed")) %>% 
  na.omit()

## Actual pRGRF
ankle_res_actual <- LOOCV_ankle_res_LMM %>% 
  select(ID, speed, pRGRF_N) %>% 
  spread(key = speed, value = pRGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_ankle
  )

back_res_actual <- LOOCV_back_res_LMM %>% 
  select(ID, speed, pRGRF_N) %>% 
  spread(key = speed, value = pRGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_back
  )

hip_res_actual <- LOOCV_hip_res_LMM %>% 
  select(ID, speed, pRGRF_N) %>% 
  spread(key = speed, value = pRGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_hip
  )

res_actual_df <- ankle_res_actual %>% 
  full_join(back_res_actual, by = c("ID", "speed")) %>% 
  full_join(hip_res_actual, by = c("ID", "speed")) %>% 
  na.omit()

res_actual_df <- res_actual_df %>% 
  mutate(actual = (actual_ankle + actual_back + actual_hip) / 3) %>% 
  select(ID, speed, actual) 

## Merge predicted and actual data frames
res_ANOVA_df <- res_pred_df %>% 
  full_join(res_actual_df, by = c("ID", "speed")) %>% 
  gather(
    ankle, back, hip, actual,
    key = "group",
    value = "pRGRF"
  )
res_ANOVA_df$ID    <- as.factor(res_ANOVA_df$ID)
res_ANOVA_df$speed <- as.factor(res_ANOVA_df$speed)
res_ANOVA_df$group <- as.factor(res_ANOVA_df$group)
```

```{r build_pVGRF_df, echo=FALSE}
## Predicted pVGRF
ankle_vert_pred <- LOOCV_ankle_vert_LMM %>% 
  select(ID, speed, pVGRF_N_predicted) %>% 
  spread(key = speed, value = pVGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = ankle
  )

back_vert_pred <- LOOCV_back_vert_LMM %>% 
  select(ID, speed, pVGRF_N_predicted) %>% 
  spread(key = speed, value = pVGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = back
  )

hip_vert_pred <- LOOCV_hip_vert_LMM %>% 
  select(ID, speed, pVGRF_N_predicted) %>% 
  spread(key = speed, value = pVGRF_N_predicted) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = hip
  )

vert_pred_df <- ankle_vert_pred %>% 
  full_join(back_vert_pred, by = c("ID", "speed")) %>% 
  full_join(hip_vert_pred, by = c("ID", "speed")) %>% 
  na.omit()

## Actual pVGRF
ankle_vert_actual <- LOOCV_ankle_vert_LMM %>% 
  select(ID, speed, pVGRF_N) %>% 
  spread(key = speed, value = pVGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_ankle
  )

back_vert_actual <- LOOCV_back_vert_LMM %>% 
  select(ID, speed, pVGRF_N) %>% 
  spread(key = speed, value = pVGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_back
  )

hip_vert_actual <- LOOCV_hip_vert_LMM %>% 
  select(ID, speed, pVGRF_N) %>% 
  spread(key = speed, value = pVGRF_N) %>% 
  na.omit() %>% 
  gather(
    `2`, `3`, `4`, `5`, `6`,
    key = speed,
    value = actual_hip
  )

vert_actual_df <- ankle_vert_actual %>% 
  full_join(back_vert_actual, by = c("ID", "speed")) %>% 
  full_join(hip_vert_actual, by = c("ID", "speed")) %>% 
  na.omit()

vert_actual_df <- vert_actual_df %>% 
  mutate(actual = (actual_ankle + actual_back + actual_hip) / 3) %>% 
  select(ID, speed, actual) 

## Merge predicted and actual data frames
vert_ANOVA_df <- vert_pred_df %>% 
  full_join(vert_actual_df, by = c("ID", "speed")) %>% 
  gather(
    ankle, back, hip, actual,
    key = "group",
    value = "pVGRF"
  )
vert_ANOVA_df$ID    <- as.factor(vert_ANOVA_df$ID)
vert_ANOVA_df$speed <- as.factor(vert_ANOVA_df$speed)
vert_ANOVA_df$group <- as.factor(vert_ANOVA_df$group)
```

Repeated measures ANOVAs were performed using the `ezANOVA()` function of the **ez** [package](https://cran.r-project.org/web/packages/ez/ez.pdf). Summary statistics for pRGRF and pVGRF ANOVA can be seen below, respectively.

```{r ANOVA}
# For resultant peak ground reaction force
res_ANOVA <- ezANOVA(
  data     = res_ANOVA_df,
  dv       = pRGRF,
  wid      = ID,
  within   = .(speed, group),
  detailed = TRUE,
  type     = 3
)
res_ANOVA

# For vertical peak ground reaction force
vert_ANOVA <- ezANOVA(
  data     = vert_ANOVA_df,
  dv       = pVGRF,
  wid      = ID,
  within   = .(speed, group),
  detailed = TRUE,
  type     = 3
)
vert_ANOVA
```

For either resultant and its vertical component, repeated measures ANOVA demonstrated that actual and predicted pGRF increased significantly (*p* < 0.001) along with increments in walking speed. 

<br>

To assess for any significant differences between actual and predicted pGRF in each speed, five separate repeated measures ANOVA were done, one for each walking speed.

```{r}
# For resultant peak ground reaction force
# 2 km/h
res_ANOVA_s2 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 2),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 3 km/h
res_ANOVA_s3 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 3),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 4 km/h
res_ANOVA_s4 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 4),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 5 km/h
res_ANOVA_s5 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 5),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 6 km/h
res_ANOVA_s6 <- ezANOVA(
  data     = res_ANOVA_df %>% filter(speed == 6),
  dv       = pRGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# For vertical peak ground reaction force
# 2 km/h
vert_ANOVA_s2 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 2),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 3 km/h
vert_ANOVA_s3 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 3),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 4 km/h
vert_ANOVA_s4 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 4),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 5 km/h
vert_ANOVA_s5 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 5),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)

# 6 km/h
vert_ANOVA_s6 <- ezANOVA(
  data     = vert_ANOVA_df %>% filter(speed == 6),
  dv       = pVGRF,
  wid      = ID,
  within   = group,
  detailed = TRUE,
  type     = 3
)
```

No significant differences (*p* > 0.05) between actual and predicted pGRF in each speed were found.

These results can be seen in figure below.

```{r ANOVA_plot, fig.height=5, fig.width=10, dev='svg'}
source(here("figs", "fig2.R"))
GRF_plot_grid
```

# R session info

```{r session_info}
devtools::session_info()
```

# References {#refs}